---
title: "Pod 3 - R Notebook and Resources"
author: "Audrey Keleman"
date: "11/07/2023"
output:
  html_document: 
    toc: true
    toc_float: true
---

# R Notebook Overview / Review

-   Header
    +   Specify output or "knit to.." (pdf, html, doc, ppt slides)
    +   Floating table of contents specified in this example
-   Chunks
    +   Click green button or can type in (watch out for counting ` marks)
    +   {r DESCRIPTION, arguments for this chunk}
    +   include=FALSE if don't want to include in the output
    +   message = FALSE, warning  = FALSE if you don't want messages or warnings to appear in output
    +   echo = FALSE if you don't want the CODE in the chunk to appear in output, but you want any tables / figures/ output that results from the code to appear in the output (= TRUE if you want the code to appear with the output that results from it)                
    +   eval = FALSE if you don't want the chunk to run at all when knitting (want it to be ignored)                   
    +   can specify by each chunk, and also can set specifications that apply to whole Rmd file (see example - 1st R code chunk in this file; message and warnings set to FALSE)
-   Outside chucks
    +   Text formatting outside of code chunks
    +   *italics*
    +   **bold**
    +   "-" or "*" for bulleted list; indent and "+" for sub-bullets
    +   tab and 1. numbers for numbered list
    +   extra spaces at end of a line to make a line break
    +   "#" signs for organization of headers and table of contents
-   There is an R cheat sheet for R Markdown and plenty of other fancy things online                           
  
--------------------------
Here, I find it helpful to describe the project I'm working on (brief background, aims, hypotheses); then describe dataset and sort of talk through each section through these sections with code and output to illustrate with the data.

## Preliminary settings

```{r prelims, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# this sets defaults for the entire output- these settings are to keep output cleaner;
# message = FALSE means that any messages that come up when running code will NOT appear in the output
# warning = FALSE means that any warnings that come up when running code will NOT appear in the output
```

## Setup and loading data

```{r setup, include=FALSE}

#rm(list = ls(all = TRUE))  

library(tidyverse)
library(psych)
library(data.table)
library(sjPlot) 
library(sjmisc) 
library(sjlabelled) 

library(fosdata) # data sets
```

```{r data, echo=FALSE}
wrist <- fosdata::wrist # loading data
?wrist # explanation of dataset and variables
```

  
# Miscellaneous Code to Share 

Code that has been helpful for me in ReproRehab so far!

**The data content of these is not statistically sound/content is not meaningful, just giving examples!**

## Tables - gtsummary

### Demographics table by group

```{r table, echo=FALSE}

library(gtsummary)

# level naming for categorical variables
wrist$sex <- factor(wrist$sex,
                   levels = c(0,1),
                   labels = c("Male", "Female"))

wrist$operativetreatment <- factor(wrist$operativetreatment,
                   levels = c(0,1),
                   labels = c("No Operation", "Operation"))

# creating table
wrist %>%
  subset(., select = c(age, sex, pcs_start, grip_strength_proportion, times_outside_from_home, distance_per_week, qd3m, qd12m, vas3m, vas12m, operativetreatment)) %>%
  mutate(sex = factor(sex, levels = c("Male", "Female"))) %>% 
  tbl_summary(
    missing = "no",
    by = operativetreatment,
    type = list(
      c(age, pcs_start, grip_strength_proportion, times_outside_from_home, distance_per_week, qd3m, qd12m, vas3m, vas12m) ~ "continuous",
      c(sex) ~ "categorical"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)"),
    label = list(
      age ~ "Age",
      sex ~ "Sex",
      pcs_start ~ "PCS, Start",
      grip_strength_proportion ~ "Grip Strength, Proportion",
      times_outside_from_home ~ "Times Outside Home",
      distance_per_week ~ "Distance Walked, km",
      qd3m ~ "Quick-DASH, 3 months",
      qd12m ~ "Quick-DASH, 12 months",
      vas3m ~ "VAS for Pain, 3 months",
      vas12m ~ "VAS for Pain, 12 months",
      operativetreatment ~ "Operative Treatment"
    )
  ) %>%
  add_p(
    test = list(all_continuous() ~ "t.test", all_categorical() ~ "chisq.test"),
    pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
  add_n()
```

### Regression tables

**The content of these (as with everything else here) is not statistically sound/content is not meaningful, just giving examples!**

```{r regression models, echo=FALSE}

mod1 <- lm(vas12m ~ age + sex + pcs_start + operativetreatment + grip_strength_proportion + times_outside_from_home + distance_per_week, data = wrist)
tab_model(mod1,show.aic = T, title = "DV = VAS, at 12 months")


mod2 <- lm(vas12m ~ age + sex + pcs_start + operativetreatment + grip_strength_proportion, data = wrist)

mod3 <- lm(vas12m ~ age + sex + pcs_start + operativetreatment + times_outside_from_home, data = wrist)
  
mod4 <- lm(vas12m ~ age + sex + pcs_start + operativetreatment + distance_per_week, data = wrist)

tab_model(mod2,mod3,mod4,show.aic = T, title = "DV = VAS, at 12 months", dv.labels = c("Grip Strength (Proportion)", "Times Outside the Home", "Distance Walked Per Week"))
# still nothing predicting when separate, although perhaps noteworthy is all biomarker relationships in expected direction
```


## Correlation plot 

```{r correlation plot, echo=FALSE}
library(corrplot)
# if you want stars indicating significance to appear in the corrplot figure, you need to do this step to make stars appear above R values in figure and not behind them (thus obscuring both R value and the stars indicating significance) (or use - to appear below, can change # +0.25 also based on preference)

# run this and a window will appear:     trace(corrplot, edit=TRUE)

# in that window, replace on line 443:
#place_points = function(sig.locs, point) {
#  text(pos.pNew[, 1][sig.locs], pos.pNew[, 2][sig.locs], 
#       labels = point, col = pch.col, cex = pch.cex, 
#       lwd = 2)

# with:

#place_points = function(sig.locs, point) {
#      text(pos.pNew[, 1][sig.locs], (pos.pNew[, 2][sig.locs])+0.25, 
#           labels = point, col = pch.col, cex = pch.cex, 
#           lwd = 2)

corrplot1 <- subset(wrist, select = c("age", "pcs_start", "grip_strength_proportion", "times_outside_from_home", "distance_per_week", "qd3m", "qd12m", "vas3m", "vas12m")) %>%
  drop_na() %>%
  setnames(., old = c("age", "pcs_start", "grip_strength_proportion", "times_outside_from_home", "distance_per_week", "qd3m", "qd12m", "vas3m", "vas12m"), new = c("Age", "PCS, Start of Follow-Up", "Grip Strength (Proportion of)", "Times Outside Home", "Distance (km) walked, average", "Quick-DASH, 3 months", "Quick-DASH, 12 months", "VAS for pain, 3 months", "WAS for pain, 12 months"))

res1 <- cor.mtest(corrplot1, conf.level = .95)

corrplot(cor(corrplot1),
         method = "circle",
         type = "upper",
         p.mat = res1$p,
         insig = "label_sig", #blank
         sig.level = c(.001, .01, .05),
         pch.cex = 0.6,
         number.cex = 0.5,
         #pch.col = "black",
         tl.col = "black",
         #tl.cex = 1,
         addCoef.col = "black",
         #tl.pos = "n",
         outline = FALSE,
         diag = FALSE)
```


## Matching two data frames by nearest date and participant ID

```{r match by nearest date, eval=FALSE, include=FALSE}

# loading data
pass_dates_yr1 <- read.csv("/Users/audreykeleman/Library/CloudStorage/Box-Box/00 keleman/projects/Falls_Preclinical_AD/FallsAndPreclinicalA_DATA_2023-09-19_0957.csv")
csf <- read.csv("/Users/audreykeleman/Library/CloudStorage/Box-Box/00 keleman/projects/Falls_Preclinical_AD/csf_markers.csv")

# assigning IDs as factors
pass_dates_yr1$id <- as.factor(pass_dates_yr1$id)
csf$id <- as.factor(csf$id)

# assigning dates as dates
pass_dates_yr1$visitdate <-as.Date(pass_dates_yr1$visitdate, format = "%m/%d/%y") 
csf$CSF_LP_DATE <-as.Date(csf$CSF_LP_DATE, format = "%m/%d/%y") 

# creating the function
MatchbyNearestDate<-function(df1, df2, ID, Date1, Date2){
  z <- lapply(intersect(df1[,ID],df2[,ID]),function(id) {
    df1 <- df1[df1[,ID] == id,]
    df2 <- df2[df2[,ID] == id,]
    
    df1[,"indices"] <- sapply(df1[,Date1],function(d) which.min(abs(df2[,Date2] - d)))
    df2[,"indices"] <- 1:nrow(df2)
    
    merge(df1,df2,by=c(ID,'indices'))
  })

  df_matched <- do.call(rbind,z)
 df_matched$indices <- NULL
  return(df_matched)
  }

# using the function
csf_yr1<-MatchbyNearestDate(pass_dates_yr1, csf, "id", "visitdate", "CSF_LP_DATE")
```

This code creates a function (called MatchbyNearestDate) that takes two data frames (df1 and df2), filters them based on common IDs, finds the nearest date matches within each group (each common ID for each data frame- date in df2 closest to that in df1), and then merges the matched rows into a single data frame, which is returned as the output.
